GPU Efficient Texture Based Bezier Curve Evaluation

### 2015-08-12 Paper/paper.tex
Started working on feedback from paper

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}
There are two basic ways for a shader program to have customized curve data.  One way is to evaluate the curve on the CPU at discrete intervals and put those points into a texture that can be used by the GPU.  The other way is to pass curve control point data from the CPU to the shader program as shader constants, allowing the shader program to evaluate the curve at a given time $t$.

@@

\begin{figure}
  \includegraphics[width=5in]{Figure2.png}
  \caption{Quadratic Bezier curves (yellow) compared to a 4x1 baked out (quantized) texture for the same curve (red).  \textit{Left:} Curve evaluated from shader program constants. \textit{Right:}  The technique from this paper - the curve is stored in a 2x2 texture and a bilinear texture sample is taken to get each point on the curve.}
  \label{fig:quickcomparison}
\end{figure}

@@

The De Casteljeau algorithm for evaluating a quadratic curve (degree 2) is also just a linear interpolation between two linear interpolations.  If the control points of the quadratic curve are $P_0,P_1,P_2$, let's call the first interpolated point $Q_{0,1}$ which is the interpolation between $P_0$ and $P_1$ for time $t$. The second linear interpolation is $Q_{1,2}$ and is the interpolation between $P_1$ and $P_2$ for time $t$.  The final point on the curve will be the interpolation at time $t$ between $Q_{0,1}$ and $Q_{1,2}$, which we can call $R_{0,1,2}$.

We could set up a texture such that the two $x$ axis interpolations would give us the values $Q_{0,1}$ and $Q_{1,2}$ respectively, and then the $y$ axis interpolation between those values would give us $R_{0,1,2}$.  When sampling, we would also use texture coordinates such it results in the same time $t$ value for interpolation on each axis.

@@

Knowing that a quadratic curve is just the linear interpolation of time $t$ between the points on two linear curves at time $t$, we have to find a way to make a texture that will cause this to happen with bilinear interpolation.  The result is that if we have a $2$x$2$ texture, the top two pixels should be $P_0$ and $P_1$ and the bottom two pixels should be $P_1$ and $P_2$.  When we do a bilinear interpolation such that we plug $t$ into both $u$ and $v$, the end result will be that the interpolations between $\overline{P_0P_1}$ and $\overline{P_1P_2}$ will be calculated from the $u$ axis interpolations, and then those two will be linearly interpolated across the $v$ axis to give us the final result.

@@

Knowing that a cubic curve is just the linear interpolation of time $t$ between the points on two quadratic curves at time $t$, we need to find a way to set up a 3d texture to make this happen with trilinear texture sampling.  With a $2$x$2$x$2$ texture, the front $2$x$2$ pixels would encode a quadratic curve for $P_0,P_1,P_2$ and the back $2$x$2$ pixels would encode a quadratic curve for $P_1,P_2,P_3$.  Then, when we sample the texture, plugging $t$ into $u,v$ and $w$, the result will be that it will calculate the points on the quadratic curves $P_0,P_1,P_2$ at time $t$ and also $P_1,P_2,P_3$ at time $t$, and then will interpolate across the $w$ axis to give us the final result.

@@

Evaluating a piecewise curve with this technique is very simple.  As an example, let's say that we have a texture that is 8x2, and encodes 4 quadratic Bezier curves end to end. The left most 2x2 pixels encode the quadratic curve used for time $t \epsilon [0,0.25)$.  The next block of 2x2 pixels encode the quadratic curve used for time $t \epsilon [0.25,0.5)$.  The third block encodes time $t \epsilon [0.5,0.75)$, and the last block encodes time $t \epsilon [0.75,1.0]$.

@@

This technique does have applications to vector graphics, but they are pretty limited and there are better solutions if using Bezier curves as vector art is the goal \cite{Loop:2005:RIC:1073204.1073303,Nehab:2008:RRG:1409060.1409088}, or if the goal is to use textures for vector art data storage \cite{Green:2007:IAM:1281500.1281665}.  The limitation is due to the fact that this technique is good at calculating the point on a curve at time $t$, but when using curves in vector graphics situations, you usually do not know the time $t$ that you want to know the curve point for, and it is usually difficult to find it.

However, there are certain special situations where you can easily get the time $t$ value of the curve that you care about.  Two of the most common situations for these are when rendering 1D explicit Bezier curves, and when rendering 1D explicit Bezier curves in polar form.

@@

Extending this technique to the third dimension would be an interesting usage case, as well as exploring ways to make the vector art usage cases more attractive.

>>>>

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}
There are two basic ways for a shader program to have customized curve data.  One way is to evaluate the curve on the CPU at discrete intervals and put those points into a texture that can be used by the GPU.  The other way is to pass curve control point data from the CPU to the shader program as shader constants, allowing the shader program to evaluate the curve at a given $t$.

@@

\begin{figure}
  \includegraphics[width=5in]{Figure2.png}
  \caption{Quadratic Curve  \textit{Left:} Yellow is the curve calculated from shader program constants.  Red is the same curve baked out in a 4x1 texture.  \textit{Right:}  Yellow uses the technique from this paper to encode the curve in a 2x2 texture.  Red is the same curve baked out in a 4x1 texture again for comparison.}
  \label{fig:quickcomparison}
\end{figure}

@@

The De Casteljeau algorithm for evaluating a quadratic curve (degree 2) is also just a linear interpolation between two linear interpolations.  If the control points of the quadratic curve are $P_0,P_1,P_2$, let's call the first interpolated point $Q_{0,1}$ which is the interpolation between $P_0$ and $P_1$ at $t$. The second linear interpolation is $Q_{1,2}$ and is the interpolation between $P_1$ and $P_2$ at $t$.  The final point on the curve will be the interpolation at $t$ between $Q_{0,1}$ and $Q_{1,2}$, which we can call $R_{0,1,2}$.

We could set up a texture such that the two $x$ axis interpolations would give us the values $Q_{0,1}$ and $Q_{1,2}$ respectively, and then the $y$ axis interpolation between those values would give us $R_{0,1,2}$.  When sampling, we would also use texture coordinates such it results in the same $t$ value for interpolation on each axis.

@@

Knowing that a quadratic curve is just the linear interpolation at $t$ between the points on two linear curves at $t$, we have to find a way to make a texture that will cause this to happen with bilinear interpolation.  The result is that if we have a $2$x$2$ texture, the top two pixels should be $P_0$ and $P_1$ and the bottom two pixels should be $P_1$ and $P_2$.  When we do a bilinear interpolation such that we plug $t$ into both $u$ and $v$, the end result will be that the interpolations between $\overline{P_0P_1}$ and $\overline{P_1P_2}$ will be calculated from the $u$ axis interpolations, and then those two will be linearly interpolated across the $v$ axis to give us the final result.

@@

Knowing that a cubic curve is just the linear interpolation of $t$ between the points on two quadratic curves at $t$, we need to find a way to set up a 3d texture to make this happen with trilinear texture sampling.  With a $2$x$2$x$2$ texture, the front $2$x$2$ pixels would encode a quadratic curve for $P_0,P_1,P_2$ and the back $2$x$2$ pixels would encode a quadratic curve for $P_1,P_2,P_3$.  Then, when we sample the texture, plugging $t$ into $u,v$ and $w$, the result will be that it will calculate the points on the quadratic curves $P_0,P_1,P_2$ at $t$ and also $P_1,P_2,P_3$ at $t$, and then will interpolate across the $w$ axis to give us the final result.

@@

Evaluating a piecewise curve with this technique is very simple.  As an example, let's say that we have a texture that is 8x2, and encodes 4 quadratic Bezier curves end to end. The left most 2x2 pixels encode the quadratic curve used at $t \epsilon [0,0.25)$.  The next block of 2x2 pixels encode the quadratic curve used at $t \epsilon [0.25,0.5)$.  The third block encodes $t \epsilon [0.5,0.75)$, and the last block encodes $t \epsilon [0.75,1.0]$.

@@

This technique does have applications to vector graphics, but they are pretty limited and there are better solutions if using Bezier curves as vector art is the goal \cite{Loop:2005:RIC:1073204.1073303,Nehab:2008:RRG:1409060.1409088}, or if the goal is to use textures for vector art data storage \cite{Green:2007:IAM:1281500.1281665}.  The limitation is due to the fact that this technique is good at calculating the point on a curve at $t$, but when using curves in vector graphics situations, you usually do not know the $t$ that you want to know the curve point for, and it is usually difficult to find it.

However, there are certain special situations where you can easily get the $t$ value of the curve that you care about.  Two of the most common situations for these are when rendering 1D explicit Bezier curves, and when rendering 1D explicit Bezier curves in polar form.

@@

Extending this technique to the third dimension to make surfaces would be an interesting usage case, as well as exploring ways to make the vector art usage cases more attractive.
